---
title: "Topic Model Analysis"
output: html_notebook
---

```{r}
library(tidyverse)
low_color <- '#660066'
high_color <- '#ffffff'
```

## Coherence

```{r}
data <- read_tsv("results/coherence.tsv") %>%
  mutate(
    subtask=subtask %>%
      str_replace(fixed(".sub.lower"), ".none") %>%
      str_remove(fixed(".sub")) %>%
      str_remove(fixed(".lower")) %>%
      str_remove(fixed(".parsed")) %>%
      str_split(fixed("."))) %>%
  hoist(
    subtask,
    language=1,
    lemmatization=2,
    topic_model=3,
    stop_word_filtering=4) %>%
  mutate(
    lemmatization=lemmatization %>% str_remove("^lem-"),
    stop_word_filtering=stop_word_filtering %>% str_remove("^stop-"),
    topic_model=topic_model %>%
      str_remove("^topic-model-") %>%
      str_split(fixed("-"))) %>%
  hoist(topic_model, num_topics=1, trial=2) %>%
  type_convert(col_types=cols(
    language=col_factor(levels=c("en", "fa", "ko", "ru")),
    lemmatization=col_factor(levels=c("none", "treetagger", "udpipe")),
    num_topics=col_integer(),
    trial=col_factor(levels=as.character(0:9)),
    stop_word_filtering=col_factor(),
    score=col_double()))
data %>%
  mutate(score=-score) %>%
  ggplot(aes(x=lemmatization, y=score)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=-45, hjust=0)) +
  facet_grid(~ language) +
  ylab("negative coherence")
ggsave('results/coherence.pdf', width=5, height=4)
```

### Sanity Check

```{r}
data_sanity <- read_tsv("results/coherence-sanity.tsv") %>%
  mutate(
    subtask=subtask %>%
      str_replace(fixed(".sub.lower"), ".none") %>%
      str_remove(fixed(".sub")) %>%
      str_remove(fixed(".lower")) %>%
      str_remove(fixed(".parsed")) %>%
      str_split(fixed("."))) %>%
  hoist(
    subtask,
    language=1,
    lemmatization=2,
    topic_model=3,
    stop_word_filtering=4) %>%
  mutate(
    lemmatization=lemmatization %>% str_remove("^lem-"),
    stop_word_filtering=stop_word_filtering %>% str_remove("^stop-"),
    topic_model=topic_model %>%
      str_remove("^topic-model-") %>%
      str_split(fixed("-"))) %>%
  hoist(topic_model, num_topics=1, trial=2) %>%
  type_convert(col_types=cols(
    language=col_factor(levels=c("en", "fa", "ko", "ru")),
    lemmatization=col_factor(levels=c("none", "treetagger", "udpipe")),
    num_topics=col_integer(),
    trial=col_factor(levels=as.character(0:9)),
    stop_word_filtering=col_factor(),
    score=col_double()))
data %>%
  inner_join(
    data_sanity,
    by=c(
      'language',
      'lemmatization',
      'num_topics',
      'trial',
      'stop_word_filtering')) %>%
  gather(key=evaluation, value=score, score.x, score.y) %>%
  mutate(evaluation=evaluation %>% str_replace(fixed("score.x"), "generalized")) %>%
  mutate(evaluation=evaluation %>% str_replace(fixed("score.y"), "standard")) %>%
  mutate(score=-score) %>%
  ggplot(aes(x=trial, y=score, group=evaluation, fill=evaluation)) +
  geom_bar(stat='identity', position='dodge') +
  theme_bw() +
  facet_grid(lemmatization ~ language) +
  ylab("negative coherence")
```
### Unmodified Coherence

```{r}
data_sanity %>%
  mutate(score=-score) %>%
  ggplot(aes(x=lemmatization, y=score)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x=element_text(angle=-45, hjust=0)) +
  facet_grid(~ language) +
  ylab("negative coherence")
ggsave('results/coherence-unmodified.pdf', width=5, height=4)
```

## Variation of Information

```{r}
data <- read_tsv("results/voi.tsv") %>%
  mutate(
    subtask=subtask %>%
      str_replace_all(fixed(".sub.lower"), ".none") %>%
      str_remove_all(fixed(".sub")) %>%
      str_remove_all(fixed(".lower")) %>%
      str_remove_all(fixed(".parsed")) %>%
      str_split(fixed("."))) %>%
  hoist(
    subtask,
    language=1,
    lemmatization_1=2,
    topic_model_1=3,
    lemmatization_2=4,
    topic_model_2=5) %>%
  mutate(
    lemmatization_1=lemmatization_1 %>% str_remove("^lem-"),
    lemmatization_2=lemmatization_2 %>% str_remove("^lem-"),
    topic_model_1=topic_model_1 %>%
      str_remove("^topic-model-") %>%
      str_split(fixed("-")),
    topic_model_2=topic_model_2 %>%
      str_remove("^topic-model-") %>%
      str_split(fixed("-"))) %>%
  hoist(topic_model_1, num_topics_1=1, trial_1=2) %>%
  hoist(topic_model_2, num_topics_2=1, trial_2=2) %>%
  type_convert(col_types=cols(
    language=col_factor(levels=c("en", "fa", "ko", "ru")),
    lemmatization_1=col_factor(levels=c("none", "treetagger", "udpipe")),
    lemmatization_2=col_factor(levels=c("none", "treetagger", "udpipe")),
    num_topics_1=col_integer(),
    num_topics_2=col_integer(),
    trial_1=col_integer(),
    trial_2=col_integer(),
    score=col_double())) %>%
  filter((lemmatization_1 != lemmatization_2) | (trial_1 > trial_2))
data %>% count(language, lemmatization_1, lemmatization_2)
```

```{r}
data %>%
  ggplot(aes(x=lemmatization_2, y=score)) +
  geom_boxplot(notch=T) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=-45, hjust=0)) +
  facet_grid(language ~ lemmatization_1, scales='free_y')
ggsave('results/voi-boxplot.pdf', width=5, height=8)
```

```{r}
data %>%
  ggplot(aes(x=score)) +
  geom_histogram() +
  theme_bw() +
  facet_grid(interaction(lemmatization_1, lemmatization_2) ~ language, scales='free_y')
```
### Intra-treatment VOI is lower than inter-treatment VOI

```{r}
tibble(one=1, lang=data$language %>% unique) %>%
  inner_join(tibble(one=1, lem=data$lemmatization_1 %>% unique), by='one') %>%
  inner_join(tibble(one=1, lem_other=data$lemmatization_1 %>% unique), by='one') %>%
  select(-one) %>%
  filter(lem != lem_other) %>%
  rowwise %>%
  mutate(p_value=t.test(
    (data %>% filter(language == lang, lemmatization_1 == lem, lemmatization_2 == lem))$score,
    (data %>% filter(language == lang, lemmatization_1 == lem, lemmatization_2 == lem_other))$score,
    alternative='less'
  )$p.value)
```
### Korean and Russian: Lemmatized treatments are closer to each other than to the unlemmatized treatment

```{r}
t.test(
  (data %>% filter(language == 'ko', lemmatization_1 == 'treetagger', lemmatization_2 == 'udpipe'))$score,
  (data %>% filter(language == 'ko', lemmatization_1 == 'treetagger', lemmatization_2 == 'none'))$score,
  alternative='less'
)
t.test(
  (data %>% filter(language == 'ko', lemmatization_1 == 'treetagger', lemmatization_2 == 'udpipe'))$score,
  (data %>% filter(language == 'ko', lemmatization_1 == 'udpipe', lemmatization_2 == 'none'))$score,
  alternative='less'
)
t.test(
  (data %>% filter(language == 'ru', lemmatization_1 == 'treetagger', lemmatization_2 == 'udpipe'))$score,
  (data %>% filter(language == 'ru', lemmatization_1 == 'treetagger', lemmatization_2 == 'none'))$score,
  alternative='less'
)
t.test(
  (data %>% filter(language == 'ru', lemmatization_1 == 'treetagger', lemmatization_2 == 'udpipe'))$score,
  (data %>% filter(language == 'ru', lemmatization_1 == 'udpipe', lemmatization_2 == 'none'))$score,
  alternative='less'
)
```

```{r}
t.test(
  (data %>% filter(language == 'en', lemmatization_1 == 'treetagger', lemmatization_2 == 'udpipe'))$score,
  (data %>% filter(language == 'en', lemmatization_1 == 'udpipe', lemmatization_2 == 'none'))$score,
  alternative='greater'
)
t.test(
  (data %>% filter(language == 'fa', lemmatization_1 == 'treetagger', lemmatization_2 == 'udpipe'))$score,
  (data %>% filter(language == 'fa', lemmatization_1 == 'treetagger', lemmatization_2 == 'none'))$score,
  alternative='greater'
)
```

### Lemmatized intra-treatment VOI is greater than unlemmatized in most cases

```{r}
tibble(one=1, lang=data$language %>% unique) %>%
  inner_join(tibble(one=1, lem=data$lemmatization_1 %>% unique), by='one') %>%
  select(-one) %>%
  filter(lem != 'none') %>%
  rowwise %>%
  mutate(p_value=t.test(
    (data %>% filter(language == lang, lemmatization_1 == lem, lemmatization_2 == lem))$score,
    (data %>% filter(language == lang, lemmatization_1 == 'none', lemmatization_2 == 'none'))$score,
    alternative='greater'
  )$p.value)
```

```{r}
t.test(
  (data %>% filter(language == 'ko', lemmatization_1 == 'udpipe', lemmatization_2 == 'udpipe'))$score,
  (data %>% filter(language == 'ko', lemmatization_1 == 'none', lemmatization_2 == 'none'))$score,
  alternative='less'
)
```

### Intra-treatment VOI on Farsi is greater for UDPipe than for TreeTagger

```{r}
t.test(
  (data %>% filter(language == 'fa', lemmatization_1 == 'udpipe', lemmatization_2 == 'udpipe'))$score,
  (data %>% filter(language == 'fa', lemmatization_1 == 'treetagger', lemmatization_2 == 'treetagger'))$score,
  alternative='greater'
)
```

### Summary plots

```{r}
data_summarized <- data %>%
  group_by(language, lemmatization_1, lemmatization_2) %>%
  summarize(`mean VOI`=mean(score)) %>%
  ungroup
data_summarized %>%
  ggplot(aes(x=lemmatization_1, y=lemmatization_2, fill=`mean VOI`)) +
  geom_tile() +
  scale_fill_gradient(high=high_color, low=low_color) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=-45, hjust=0)) +
  xlab("lemmatization") +
  ylab("lemmatization") +
  facet_wrap(~ language)
ggsave('results/voi.pdf', width=5, height=4)
```


```{r}
for (lang in c('en', 'fa', 'ko', 'ru')) {
  data_summarized %>%
    filter(language == lang) %>%
    ggplot(aes(x=lemmatization_1, y=lemmatization_2, fill=`mean VOI`)) +
    geom_tile() +
    scale_fill_gradient(high=high_color, low=low_color) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=-45, hjust=0)) +
    xlab("") +
    ylab("")
  ggsave(str_c('results/voi-', lang, '.pdf'), width=3.5, height=2.5)
}
```